maxent model avec xvalidation (year-based test training)

### 2019-04-12 src/maxent_avec_xvalidation.Rmd
start listing reg values needed
minor updates to code for march

```{r pre-install & load packages, include=FALSE}
packages <- c("dismo", "stringr", "rJava")
source("../src/ipak.R")
ipak(packages)
```

@@

jan2003: b = 1
jan2004: b = 1
jan2005: b = 1.5
jan2006: b = 2.25

@@

  if (timeslice == "jan2003"){
    betaval <- 1
  } else if (timeslice == "jan2004"){
    betaval <- 1
  } else if (timeslice == "jan2005"){
    betaval <- 1.5
  } else if (timeslice == "jan2006"){
    betaval <- 2.25
  }

@@

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
    stackdeplst[[year]] <- stackdep
    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

    for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, chl_surface, o2_depth, salinity_depth, nao_sample, nao_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
      dev.off() # stops automatic saving of the plot to a png
    }
}

>>>>

```{r pre-install & load packages, include=FALSE}
options(java.parameters = "-Xmx1g" )#needs to run before dismo loaded - boosts memory for java to 1g
packages <- c("dismo", "stringr", "rJava")
source("../src/ipak.R")
ipak(packages)
```

@@

KEEP A NOTE OF THESE IN ONENOTE

mar2003 b = 1

mar1999 b = 1.25
mar2002 b = 1.25
mar2008 b = 1.25
mar2009 b = 1.25

mar2005 b = 1.5
mar2006 b = 1.5

mar2000 b = 1.75

mar2001 b= 2.25
mar2007 b = 2.25

mar2010 b = 3.5

@@

  if (timeslice == "mar2003"){
    betaval <- 1
  } else if (timeslice == "mar1999"| timeslice == "mar2002" | timeslice == "mar2008" | timeslice == "mar2009"){
    betaval <- 1.25
  } else if (timeslice == "mar2005" | timeslice == "mar2006"){
    betaval <- 1.5
  } else if (timeslice == "mar2000"){
    betaval <- 1.75
  } else if (timeslice == "mar2001" | timeslice == "mar2007"){
    betaval <- 2.25
  } else if (timeslice == "mar2010"){
    betaval <- 3.5
  }

@@

    unlayer_yr <- subset(unlayer, ascyears == yr)
    unlayer_yr <- unlayer_yr[-c(1), ]
    unlayer_yr[] <- lapply(unlayer_yr, as.character) #from https://stackoverflow.com/questions/2851015/convert-data-frame-columns-from-factors-to-characters
    stackdeplst <- list()
    stackdep <- subset(unlayer_yr, select = c(stackname, ascdepths))
    stackdeplst[[year]] <- stackdep
    #write.csv(stackdep, paste0("../output/maxent/predictions/stackdepthlist_", yr, ".csv"), row.names = FALSE)
    aea <- raster("../output/env/aea.tif")

    for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_prev <- raster(unlayer_yr$nao_prev[td])
      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_prev) <- "nao_prev"
      nao_prev <- projectRaster(nao_prev, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      amo_sample <- raster(unlayer_yr$amo_sample[td])
      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(amo_sample) <- "amo_sample"
      amo_sample <- projectRaster(amo_sample, aea)
      amo_winter <- raster(unlayer_yr$amo_winter[td])
      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(amo_winter) <- "amo_winter"
      amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/predictions/", mthna, "/", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      prplot <- plot(pred, png(paste0("../output/maxent/predictions/", mthna, "/", stackn, ".png"))) 
      dev.off() # stops automatic saving of the plot to a png
    }
}
